---
title: "Random Secret Santa R Script"
description: |
  A script I created to automate my family's yearly secret santa gift exchange
author:
  - name: Peter Menzies
    url: {}
date: 08-23-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

### My family's yearly gift exchange

Here's a little context needed to understand the program I created, before we get to the actual code. 

Each year my extended family does a secret santa gift exchange where each member of the family is randomly paired with another — the one parameter being that each person must be matched with someone outside of their nuclear family. The matches are assigned using the good old "pick a name from the hat" method. If you draw a name from within your immediate family, you try again until you get someone who isn't. This process seemed like an excellent candidate for automation, and with a couple weeks of R under my belt, it felt like a perfect task to apply what I had learned so far.

### Packages used: `tidyverse` and `here`

```{r}
library(tidyverse)
library(here)
```


### Creating data frame with family member's and their corresponding family unit

I read in a CSV that contains two variables we care about: `Person` and `Group`. Each family member is listed in the `Person` column, and the label created for their family unit under `Group`. 

```{r}
xmas_names <- read_csv(here::here("data", "xmas_names.csv")) %>% 
  select(Person, Group) 
```

### Creating the function

I want to create a self-contained function that can be run repeatedly, without affecting the original `xmas_names` data frame. Let's start by creating the temporary data frames and vectors we'll need during the matching iteration process.

```{r}
dim_names <- xmas_names
matches <- vector(mode = "character", length = 13)
already_matched <- vector(mode = "character")
```

- `dim_names` will be our list of *diminishing* names that we will iterate through and will remove names from that have already been matched. 

- `matches` is our empty vector that will be populated with strings containing 

