---
title: "User input as function arguments in R"
description: |
  A brief rundown of using readline() to prompt a user's input for use as function arguments
author:
  - name: Peter Menzies
    url: {}
date: 08-22-2021
output:
  distill::distill_article:
    self_contained: false
categories:
  - R
  - "Function I Learned"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = 'hide')
library(tidyverse)
library(palmerpenguins)
```

### Let's create a simple function that outputs a statement containing a person's name and the amount of change in their pocket given the number of each US coin they have

```{r}
spare_change <- function(friend, pennies, nickels, dimes, quarters) {

  penny_val <- pennies * 0.01
  nickel_val <- nickels * 0.05
  dime_val <- dimes * 0.10
  quarter_val <- quarters * 0.25

  all_change <- sum(penny_val, nickel_val, dime_val, quarter_val)

  return(paste("Your friend", friend, "has $", all_change))
}
```

### This will work just fine, but that's a lot of required arguments — we can make it a bit more user-friendly by prompting the user for input that will supply our arguments.

### The `readline()` function takes input from the user which we can assign to an object. Within `readline()`, we define the argument `prompt` as the question or directive we want presented to the user. For example:

```{r}
friend <- readline(prompt = "what is your friend's name?")
```

### In the above example, `friend` is now assigned the value of whatever the user enters following the prompt.

### We can use this same function to gather all the needed arguments for `spare_change()` by prompting the user for input. This will look like:

```{r}
spare_change <- function() {

  friend <- readline(prompt = "Enter a friend's name: ")
  pennies <- as.numeric(readline(prompt = "How many pennies do they have: "))
  nickels <- as.numeric(readline(prompt = "How many nickels do they have: "))
  dimes <- as.numeric(readline(prompt = "How many dimes do they have: "))
  quarters <- as.numeric(readline(prompt = "How many quarters do they have: "))

  penny_val <- pennies * 0.01
  nickel_val <- nickels * 0.05
  dime_val <- dimes * 0.10
  quarter_val <- quarters * 0.25

  all_change <- sum(penny_val, nickel_val, dime_val, quarter_val)

  return(paste("Your friend", friend, "has $", all_change))
}
```

##### It's important to note that `readline()` returns a string, so for our coin value calculations, we had to convert the user's input to a numeric value.

### Our function now requires no arguments when called — it will gather them from the user. Our user can run the function without using the documentation to determine what arguments are needed.


