---
title: "SecRet Santa"
description: |
  An R function I built to automate my family's secret santa matching process
author:
  - name: Peter Menzies
    url: {}
date: 08-23-2021
output:
  distill::distill_article:
    self_contained: false
categories:
  - R
slug: "2021-23-08-secret-santa-r-script"
preview: chRistmas_wide.png
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### A litle context...

Every year my extended family does a gift exchange where each member of the family is randomly paired with another member—the one caveat being that each person must be matched with someone outside of their immediate family. Traditionally, the matches are assigned using the good old "pick a name from a hat" method. If you draw a name from within your immediate family, you try again until you get someone who isn't. This process seemed like an excellent candidate for automation, and with a couple weeks of R under my belt it felt like a perfect project to apply what I had learned so far.

### Packages used: `tidyverse`, `here`, and `gt`

```{r}
library(tidyverse)
library(here)
library(gt)
```


### Reading in data

Firstly, I read in a CSV that contains two variables we care about: `Person` and `Group`. Each family member is listed in the `Person` column, and the label created for their family unit under `Group`. 

```{r}
xmas_names <- read_csv(here("data", "xmas_names.csv")) %>% 
  select(Person, Group) 
```

### Creating the function

I want to create a function that outputs a dataframe with two columns, `Giver` and `Recipient`, that lists each member of the family and who they will be giving a gift to. Let's call the function `match_maker`.

The function will first create the framework of the aforementioned `ouput_df` and a vector that we'll populate with the names of those that have already been matched.

```{r, eval=FALSE}
match_maker <- function() {

  output_df <- data.frame(Giver = seq(1:nrow(xmas_names)), Recipient = seq(1:nrow(xmas_names)))
  already_matched <- vector(mode = "character")
```

### For loop

Within our function we'll have a for loop that iterates through each row in our `xmas_names` dataframe and randomly pairs the `Person` in each row with another person from outside of their immediate family unit (`Group`). 

The loop will iterate along the sequence `1:nrow(xmas_names)`.

In order to adjust our pool of "allowable" matches within each iteration, we're going to create a temporary dataframe—`temp_names`—that is first assigned to `xmas_names` and then filtered based on our two criteria: the person isn't in the same `Group` and hasn't been matched already. We'll use `dplyr::filter()` to subset `temp_names` to meet these parameters.

```{r, eval=FALSE}
  for (i in 1:nrow(xmas_names)) {
    
    temp_names <- xmas_names %>%
      filter(Group != xmas_names$Group[i]) %>% 
      filter(!Person %in% already_matched)
```

Next, before the loop goes any further we'll need to add an `if` statement that saves our function from hitting a dead end. Depending on the order in which matches are randomly assigned, it's possible for only members of a single family unit to remain in the final iterations. When this happens, our `temp_names` dataframe will be empty because there are no longer any viable matches, and our code will break. The `if` statement will restart the function if `temp_names` becomes empty before all people have been matched.

```{r, eval=FALSE}
    if (nrow(temp_names) == 0) {
      return(match_maker())
```

Next we're going to generate a random number using the `sample()` function and assign it to `sample_number`. We'll use this number to index `temp_names$Person` and assign that value to `match`. `sample_number` has to be in the range `1:nrow(temp_names)` because `temp_names` is our temporary dataframe of viable match candidates. The person being matched we can call `matchee`.

And we need to `append()` our current `match` to our vector `already_matched` which is used to filter `temp_names` so that person can't be selected as a recipient again.

```{r, eval=FALSE}
    }
    
    sample_number <- sample(1:nrow(temp_names), 1)
    
    match <- temp_names$Person[sample_number]
    
    matchee <- xmas_names[[i, 1]]
    
    already_matched <- append(already_matched, match)
```

Lastly, we'll add `matchee` and `match` to the `output_df`, and instruct the function to return `output_df` once the for loop has iterated through all names on the list.

```{r, eval=FALSE}
    output_df$Giver[i] <- matchee
    output_df$Recipient[i] <- match
    
  }
  
  return(output_df)
  
}
```

### Function in full:

```{r}
match_maker <- function() {

  output_df <- data.frame(Giver = seq(1:nrow(xmas_names)), Recipient = seq(1:nrow(xmas_names)))
  already_matched <- vector(mode = "character")
  
  for (i in 1:nrow(xmas_names)) {
    
    temp_names <- xmas_names %>%
      filter(Group != xmas_names$Group[i]) %>% 
      filter(!Person %in% already_matched)
    
    if (nrow(temp_names) == 0) {
      return(match_maker())
    }
    
    sample_number <- sample(1:nrow(temp_names), 1)
    
    match <- temp_names$Person[sample_number]
    
    matchee <- xmas_names[[i, 1]]
    
    already_matched <- append(already_matched, match)
    
    output_df$Giver[i] <- matchee
    output_df$Recipient[i] <- match
    
  }
  
  return(output_df)
  
}
```

### The result:

Now let's make some matches! (I used `gt()` here just to create a nice table for the output)

```{r}
match_maker() %>% gt()
```








